### 2014-04-05 paper/abstract.tex
george comments

\begin{abstract}
3D printers offer extraordinary flexibilty for prototyping the shape and mechanical function of objects.  However, a 3D printer's role in prototyping interactive objects is still not clear.  We propose a general technique for supporting the rapid prototyping of interactivity by removing interior material of 3D models to form pipes and chambers. We describe this new design space of pipes and hollow chambers for interaction design, where variables include openings, topologies, design focus, and inserted media \george{this needs more description or explanation}.  We then present a  tool for routing such pipes through the interior of 3D models, integrated within a 3D modeling program.  We use two distinct routing algorithms.  The first allows users to define the pipes' terminals, and then uses A* path routing and physics-based simulation to route the paths with minimized bending energy, to allow easy insertion of media post-print.  The second allows users to supply a desired internal shape which we fit a pipe route to: for this we developed a novel graph-based routing algorithm.  We present several totally tubular prototypes we created using our tool to show its flexibility and potential.
\end{abstract}

>>>>

\begin{abstract}
3D printers offer extraordinary flexibilty for prototyping the shape and mechanical function of objects.  However, a 3D printer's role in prototyping interactive objects is still not clear.  We propose a general technique for supporting the rapid prototyping of interactivity by removing interior material of 3D models to form pipes and chambers. We describe this new design space of pipes and hollow chambers for interaction design, where variables include openings, topologies, design focus, and inserted media.  We then present a  tool for routing such pipes through the interior of 3D models, integrated within a 3D modeling program.  We use two distinct routing algorithms.  The first allows users to define the pipes' terminals, and then uses path routing and physics-based simulation to minimize pipe bending energy, allowing easy insertion of media post-print.  The second allows users to supply a desired internal shape to which we fit a pipe route: for this we developed a novel graph-based routing algorithm.  We present several totally tubular prototypes we created using our tool to show its flexibility and potential.
\end{abstract}

### 2014-04-05 paper/designspace.tex
george comments

\george{difference between open/return not clear}Pipes can have four opening styles: open, return, semi-closed, and fully enclosed (see Figure \ref{fig:pipespace}).  As we are interested in interactive devices, we distinguish a ``user side"---part of the surface or interior of an object facing its user for interaction; and a ``system side"---where other hardware such as electronics may be connected.  We refer to a pipe which is ``open on the user side'' to have its contents \emph{physically accessible} to a user \emph{at interaction time}.  ``Closed on the user side'' implies physically cut off from the user.  Similarly with the system side.  User or system manipulation of pipe contents are possible even without physical access, e.g., via soundwaves. \george{confusing}

@@

Pipe network topologies enable different interactions \george{how?}.  Splitting or mixing pipes offer flexibility in output, allowing a single source to have multiple outputs or multiple sources to combine (see Figure \ref{fig:pipespace}).  Star and tree topologies extend the splitting and mixing primitives, encompassing multiple-in and multiple-out in contrast to single-in/single-out; we used star topologies in our toys \ref{fig:toys}.

@@

Pipes need not have hollow centers: in the case where routed pipes are filled with solid material---in particular, a solid material different from the enclosing model material---, interactions such as those in Printed Optics \cite{Willis-printedoptics} are possible.

Particulates, either printed in-place or inserted, can be of varying densities.  A single particle can be used for display.  Sparse particles in fluid can provide haptic feedback.

>>>>

\george{difference between open/return not clear}Pipes can have four opening styles: open, return, semi-closed, and fully enclosed (see Figure \ref{fig:pipespace}).  As we are interested in interactive devices, we distinguish a ``user side"---part of the surface or interior of an object facing its user for interaction; and a ``system side"---where other hardware such as electronics may be connected.  We refer to a pipe which is ``open on the user side'' to have its contents \emph{physically accessible} to a user \emph{at interaction time}.  ``Closed on the user side'' implies physically cut off from the user.  Similarly with the system side.  User or system manipulation of pipe contents are possible even without physical access: e.g., via soundwaves. \george{confusing}

@@

Pipe network topologies enable different interactions: splitting or mixing pipes offer flexibility in output, allowing a single source to have multiple outputs or multiple sources to combine (see Figure \ref{fig:pipespace}).  Star and tree topologies extend the splitting and mixing primitives, encompassing multiple-in and multiple-out in contrast to single-in/single-out; we used star topologies in our toys \ref{fig:toys}.

@@

Pipes need not have hollow centers: if routed pipes are filled with solid material---in particular, a solid material different from the enclosing model material---, interactions such as those in Printed Optics \cite{Willis-printedoptics} are possible.

Particulates, either printed in-place or inserted, can be of varying densities and can contain particles of various and variable size.  A single large particle can be used for display.  Small particles in fluid can provide variable stiffness haptics via jamming \cite{Follmer-jamming}.

### 2014-04-05 paper/exampleobjects.tex
george comments

\subsection{Touch-sensitive Toys (open, liquid, star, exterior)}

@@

\subsection{Breathing Bunny (semi-closed, gas, exterior)}

@@

\subsection{Custom Radio (open, threadable, exterior)}

@@

\subsection{Presence-aware Pen Holder (open, solid, exterior)}

@@

\subsection{Animated Neon Sign (return, threadable, interior)}

@@

Neon art, perhaps best known for its association with Las Vegas, is traditionally made from hand-formed glass tubes containing neon gas.  The tubes light up when a current is passed through them.  For this type of art, the path of the tubes is of crucial importance, as it determines how the sign will look.  We designed a custom neon robot head which can be animated to ``talk''.  The piece features semi-open pipes: one end must remain open to connect the EL wire to its power source.  The pipes have been threaded with EL wire which is lit in sequence to create an animation.  Our sign was fabricated on the Objet in two halves, cutting through the plane of the sign itself, with clear material for the sign around the input design and opaque everywhere else.  We manually flushed support from the non-planar connecting pipes.  The EL wire was threaded post-print, and is controlled by an EL wire sequencer.

\subsection{Maze (fully enclosed, particulate, tree, interior + exterior)}

>>>>

\subsection{Touch-sensitive Toys (open, liquid, star, terminals)}

@@

\subsection{Breathing Bunny (semi-closed, gas, terminals)}

@@

\subsection{Custom Radio (open, threadable, terminals)}

@@

\subsection{Presence-aware Pen Holder (open, solid, ternimals)}

@@

\subsection{Animated Neon Sign (return, threadable, path)}

@@

Neon art, perhaps best known for its association with Las Vegas, is traditionally made from hand-formed glass tubes containing neon gas.  The tubes light up when a current is passed through them.  For this type of art, the path of the tubes is of crucial importance, as it determines how the sign will look.  We designed a custom neon robot head which can be animated to ``talk''.  The pipes have been threaded with EL wire which is lit in sequence to create an animation.  Our sign was fabricated on the Objet in two halves, cutting through the plane of the sign itself, and with pipes cut in the rear piece to allow the wire to exit backwards and hide the EL wire controls.  The EL wire was threaded post-print, and is controlled by an EL wire sequencer.

\subsection{Maze (fully enclosed, particulate, tree, terminals + path)}

### 2014-04-05 paper/introduction.tex
george comments

\begin{figure}[h]
\centering
    \includegraphics[width=3.4in]{figures/placeholder/teaser.png}
\caption{A novel neon sign (a) and a touch-sensitive brain (b) designed with our pipe tool and fabricated on a consumer-grade 3D printer.  (c) and (d) show the structure of the internal pipes generated by our tool.  \bjoern{will this be the robot or the UIST sign?} \valkyrie{arguably, the robot will be nicer-looking.  I think I'm manufacturing the UIST sign on the Makerbot.}}
\label{fig:teaser}
\end{figure}

@@

Due to the growing popularity of 3D printing and maker communities like Thingiverse\footnote{http://thingiverse.com}, recent work has looked at the issue of rapidly prototyping interactive objects. Existing approaches include Printed Optics \cite{Willis-printedoptics}, which uses optically clear material for light redirection, and Sauron \cite{Savage-sauron}, which is based on computer vision, but they are limited to visible interactions only \george{what does this mean?}; they do not give designers freedom to integrate haptic or acoustic inputs or outputs \george{huh?  why not?}.

@@

We describe a new design space of pipes and hollow chambers for the prupose of interaction design, where variables include openings, topologies, and inserted media \george{need more detail}. By exploring this design space new opportunities for adding interactivity to 3D printed objects can be realized. For example, copper material can fill the channels, to allow for standard electronic components to be easily integrated after the printout; electroluminescent (EL) wire can be threaded through to create computer-controlled visual output; or air can be pumped through to control haptic responses.  In comparison to traditional exterior electronics integration, our process allows makers to preserve aesthetics by hiding wires inside the object; additionally it offers more freedom in selecting precise surface locations for, e.g., air mediated tactile output.

\tovi{I would remove this entire paragraph} \valkyrie{I think we at least need to preserve the sentiment from the end, that we don't have any novel I/O techniques but that pipes help extend existing techniques in useful ways} \george{remove, repeats}
Pipes allow makers to build with and upon existing input/output techniques, often integrating these methods into objects they were not originally designed for.   For example, to create a touch-sensitive toy as in Figure \ref{fig:teaser}, a maker can select several locations of interest on the brain \george{?}.  After creating pipes to those locations and printing her object, she can fill the pipes with conductive paint.  For sensing, she need only connect a single wire to a shared out of all the tubes, and can distinguish them via Swept Frequency Capacitve Sensing \cite{Sato-touche}.  In this way, she has only a single active element (the microcontroller setup), but her otherwise-inert 3D printed object (the brain) has several sensing locations.  Other existing sensing and actuation approaches, such as FlyEye \cite{Wimmer-flyeye} and Jamming User Interfaces \cite{Follmer-jamming}, can also be enhanced by the use of pipes: makers utilizing such I/O strategies can locate input and output at arbitrary points on a printed 3D object's surface.  We leverage these existing techniques for sensing and actuation while our work's novelty is in \emph{internal modeling} for 3D printing and the \emph{redirection} of I/O to arbitrary locations on a device.

\tovi{I would remove this entire paragraph}  \valkyrie{if we scrap the first one, sure.  George, what do you think?} \george{remove}
Pipes can also be used for sending media on a specified route through an object; for example, neon signs are created by evacuated glass tubes formed into a specific path.  Electricity is run through the near-vacuum, creating light.  If our maker wants to build such a sign, she can substitute EL wire for the evacuated glass: she need only design a path for it to be fed through and then create pipes along that path.

Generating these pipes, however, is not trivial.  For one, most 3D modelling tools do not provide extensive tools for modelling the  {\em interior} of 3D models. Furthermore, careful consideration needs to go into the design and routing of the pipes. Pipes need maximum possible bend radius to ease the insertion of media post-print.  Independent pipes cannot intersect, as this could lead to electrical shorting or other problems.  Pipes should not pass too close to the surface: consumer-grade 3D printers do not guarantee air- or water-tightness of prints, so more than one printed layer may be required to prevent leakage.  Pipes for applications like neon signs must have a Euler tour \george{explain or ref} through them: this allows a single EL wire to light every edge on the pipe graph.  No tools currently exist for these design tasks.

Our work makes several contributions to the field of digitally fabricated interactive objects. First, we describe a design space of how pipes can be integrated into 3D models. We then offer algorithms and techniques for routing the pipes, and integrate these tools into  a 3D modelling applicaiton. The system takes into considertation the design complexities of routing interior pipes which are described above. Finally, we showcase a set of examples, enabled by our modeling tool, which allows us to explore various points in the design space \george{what design space?}. We conclude by summarizing the input and output modalities enabled by our work, and discussing limitations and areas for future exploration.

>>>>

\begin{figure}[h]
\centering
    \includegraphics[width=3.4in]{figures/placeholder/teaser.png}
\caption{A collection of interactive objects designed using our tool.  All have electronic sensing or actuation components which are routed through the \emph{interior} of the prints to preserve aesthetics.  In (a), objects manufactured on a hobbyist 3D printer.  In (b), those manufactured on industrial-grade printers.}
\label{fig:teaser}
\end{figure}

@@

Due to the growing popularity of 3D printing and maker communities like Thingiverse\footnote{http://thingiverse.com}, recent work has looked at the issue of rapidly prototyping interactive objects. Existing approaches include Printed Optics \cite{Willis-printedoptics}, which uses optically clear material for light redirection, and Sauron \cite{Savage-sauron}, which is based on computer vision, but they are limited to visible (reflecting light off fingers, mechanical motion of components) input and visual (screen-like) output; they do not give makers freedom to integrate haptic feedback or other inputs or outputs.

@@

We describe a new design space of pipes and hollow chambers for the purpose of interaction design, where variables include openings, topologies, and inserted media. By exploring this design space new opportunities for adding interactivity to 3D printed objects can be realized. For example, copper material can fill the channels, to allow for standard electronic components to be easily integrated after the printout; electroluminescent (EL) wire can be threaded through to create computer-controlled visual output; or air can be pumped through to control haptic responses.  In comparison to traditional exterior electronics integration, our process allows makers to preserve aesthetics by hiding wires inside the object; additionally it offers more freedom in selecting precise surface locations for, e.g., air mediated tactile output.

Existing sensing and actuation approaches, such as FlyEye \cite{Wimmer-flyeye}, Swept Frequency Capacitive Sensing \cite{Sato-touche}, and Jamming User Interfaces \cite{Follmer-jamming}, can be enhanced by the use of pipes: makers utilizing such I/O strategies can locate input and output at arbitrary points on a printed 3D object's surface.  We leverage these existing techniques for sensing and actuation while our work's novelty is in \emph{internal modeling} for 3D printing and the \emph{redirection} of I/O to arbitrary locations on a device.

Generating these pipes, however, is not trivial.  For one, most 3D modelling tools do not provide extensive tools for modelling the  {\em interior} of 3D models. Furthermore, careful consideration needs to go into the design and routing of the pipes. Pipes need maximum possible bend radius to ease the insertion of media post-print.  Independent pipes cannot intersect, as this could lead to electrical shorting or other problems.  Pipes should not pass too close to the surface: consumer-grade 3D printers do not guarantee air- or water-tightness of prints, so more than one printed layer may be required to prevent leakage.  Pipes for applications like neon signs must allow a single EL wire to light every edge on the pipe graph (i.e., have an Euler tour).  No tools currently exist for these design tasks.

Our work makes several contributions to the field of digitally fabricated interactive objects. First, we describe a design space of how pipes can be integrated into 3D models. We then offer algorithms and techniques for routing the pipes, and integrate these tools into  a 3D modelling applicaiton. The system takes into considertation the design complexities of routing interior pipes which are described above. Finally, we showcase a set of examples, enabled by our modeling tool, which allows us to explore various points in the design space. We conclude by summarizing the input and output modalities enabled by our work, and discussing limitations and areas for future exploration.

### 2014-04-05 paper/tubes.tex
george comments

\section{A Series of Tubes - A Tool for Pipe Design}
\bjoern{What kind of visualizations and diagrams can help explain the technical part other than photos and screenshots? The "UIST" pipe routing figure goes in this direction. Can we make similar conceptual figures, flowcharts, etc for the earlier parts?}

Pipe design is non-trivial.  Independent pipes cannot intersect lest their contents be unintentionally mixed.  Pipes' bending energy needs to be minimized to allow easy insertion of media post-print.  Pipes in general need to avoid objects' surfaces to prevent fluid leakage when printed by hobbyist-grade machines.  Finally, for designing pipes that follow a user-specified path, we must ensure certain characteristics of that path (e.g., that the path is connected).

@@

\subsection{Exterior Connection Points}

@@

Note that pipes once cut are ``outside the mesh'', i.e., pipes cut in sequence cannot intersect each other.  However, we do not currently track all pipes for global optimization: we greedily select the best routing per pipe.

Our A* pathfinding produces the \emph{shortest} path, but in many cases
we require a smoother path through which we can actually push a physical wire.
We solve this problem via physical simulation. We create a straight virtual 
wire (a 3D poly-line or \emph{discrete rod}), and then set its initial position 
as our A* path.
Running the physical simulation allows the wire to attempt to return to
a straight configuration, while satisfying its endpoint position constraints.
We also increase the length of the wire, to provide some slack in cases where 
it would otherwise bend sharply to stay inside the shape.

@@

\subsubsection{Routing}
To allow for a single path through a pipe for post-print media insertion, we need to find a semi-Eulerian graph of which a user's input graphics are a subset, and subsequently find an Euler tour on that graph.

@@

Once we have a semi-Eulerian graph, we need to create an Euler tour.  We use a weighted modification of Fleury's algorithm for finding Euler tours\footnote{\url{http://en.wikipedia.org/wiki/Eulerian_path\#Fleury.27s_algorithm}} for this.  In the classic algorithm, each selected edge is removed to create a reduced graph.  To select the next edge while at a node, the algorithm randomly chooses from all incident non-bridge edges (where a bridge edge would create disconnected non-trivial subgraphs if it were removed).  This ensures that every edge is both reachable and traversed.  In our modification, instead of randomly selecting from non-bridge paths at each node, we select the non-bridge edge which turns the least from the most recent path (i.e., we prefer to pass straight through a node, if possible).  This minimizes turns in the final artifact, which eases support material removal and assembly.  Our EL wire in Figure \ref{fig:UIST} follows an Euler tour through the input nodes.

>>>>

\section{A Series of Tubes - A Tool for Pipe Design} \george{not clear that this is the beginning of our system description}
\bjoern{What kind of visualizations and diagrams can help explain the technical part other than photos and screenshots? The "UIST" pipe routing figure goes in this direction. Can we make similar conceptual figures, flowcharts, etc for the earlier parts?}

Pipe design is non-trivial.  Independent pipes cannot intersect lest their contents be unintentionally mixed.  Pipes' bending energy needs to be minimized to allow easy insertion of media post-print.  Pipes in general need to avoid objects' surfaces to prevent fluid leakage when printed by hobbyist-grade machines.  Finally, for designing pipes that follow a user-specified path, we must ensure certain characteristics of that path (e.g., that the path is connected). \george{pipes are hard.  not sure we convinced readers by now that they are worth the effort.  maybe this has to happen more in the intro.  the example objects section is still a few pages away.}

@@

\subsection{Exterior Terminals}

@@

Pipes avoid each other in this manner: pipes once cut are ``outside the mesh'' and therefore not part of valid routes.  However, we do not currently track all pipes for global optimization: we greedily select the best routing per pipe.

Our A* pathfinding produces the \emph{shortest} path, but in many cases
we require a smoother path through which we can actually push a physical wire.
We solve this problem via physical simulation. We create a straight virtual 
wire (a 3D poly-line or \emph{discrete rod}), and then set its initial position 
as our A* path.
Running the physical simulation allows the wire to attempt to return to
a straight configuration, while satisfying the user-selected endpoint position constraints.
We also increase the length of the wire, to provide some slack in cases where 
it would otherwise bend sharply to stay inside the shape.

@@

\subsubsection{Routing} \george{condense or just point to appendix}
To allow for a single path through a pipe for post-print media insertion, we need to find a semi-Eulerian graph of which a user's input graphics are a subset, and subsequently find an Euler tour on that graph.

@@

/george{too complicated}
Once we have a semi-Eulerian graph, we need to create an Euler tour.  We use a weighted modification of Fleury's algorithm for finding Euler tours\footnote{\url{http://en.wikipedia.org/wiki/Eulerian_path\#Fleury.27s_algorithm}} for this.  In the classic algorithm, each selected edge is removed to create a reduced graph.  To select the next edge while at a node, the algorithm randomly chooses from all incident non-bridge edges (where a bridge edge would create disconnected non-trivial subgraphs if it were removed).  This ensures that every edge is both reachable and traversed.  In our modification, instead of randomly selecting from non-bridge paths at each node, we select the non-bridge edge which turns the least from the most recent path (i.e., we prefer to pass straight through a node, if possible).  This minimizes turns in the final artifact, which eases support material removal and assembly.  Our EL wire in Figure \ref{fig:UIST} follows an Euler tour through the input nodes.
